### 2023-12-08 main.tex
in progress

For the single species case,  the solution of the exclusion process with open boundaries \cite{derrida1993exact,1993JSP....72..277S} has been of crucial importance in our understanding of the structural properties of non-equilibrium state, such as long-range correlations \cite{spohn1983long} and non-local density large deviation functions \cite{derrida2007non,derrida1998exact} and current large deviations \cite{mallick2022exact,bodineau2005current}. One may argue that the results available for the exclusion process are rooted in the algebraic symmetries of the corresponding integrable spin $s=1/2$ Heisenberg chain, see e.g. \cite{schutzManyBody,SSEPReviewRagoucy} for two excellent reviews. .
%
This integrability property can be combined with duality and leads to closed-form expressions for the correlations and the non-equilibrium steady state, as shown in \cite{frassek2020eigenstates,frassek2020duality}.
% by extending the previous results of \cite{frassek2020eigenstates}.

@@

The model studied in this paper is the {\em  multispecies stirring process with open boundaries}, i.e. the multispecies analogue of the exclusion process on a generic graph as introduced in \cite{zhou2021orthogonal} with additional boundary reservoirs as proposed in \cite{vanicat2017exact}. 
This particle system is defined on a generic connected and undirected graph. At each site 
the maximal number of allowed particles is $\nu\in \mathbb{N}$. These particles can be of $N-1$ types (or species). Two sites do interact when there is an edge between them and this interaction consists in swapping the occupations of these two sites (for a precise definition of the process see Section \ref{sec1}). 
Furthermore, reservoirs are attached at each site of the graph, creating and removing particle. The process becomes integrable when the graph is a finite chain with two reservoirs attached to the first and the last sites respectively and the maximal  number of allowed particles is fixed to be $\nu=1$,  see \cite{vanicat2017exact}. {\color{blue}The choice of the boundary reservoirs introduced in \cite{vanicat2017exact} is further motivated by the fact that they allow to determine an absorbing dual process, regardless of the geometry on which the process is defined and on the maximal number of particle hosted at each site}.
%{\color{red}Explain the model here.... $N$-species (or $N-1$ plus empty?). What is the graph? say that it becomes integrable for the line }
%We refer to the model as the {\em  multispecies stirring process with open boundaries}.
%

@@

{\color{magenta} Cristian I think it is good to keep some of the following citations, especailly the one to Schutz...
 \cite{schutz2017kardar, vanicat2017exact, crampe2016integrable, finn2018matrix, floreani2022switching, casini2022uphill, casini2023density, redig2022ergodic, redig2023equilibrium, chatterjee2023multi,  franceschini2022orthogonal,floreani2023non,carinci2023open}
See also \cite{crampe2016integrable, finn2018matrix} for the asymmetric setup.}

@@

At each vertex site $x\in V$, we describe the occupation with an $N-$dimensional vector $n^{x}=(n_{1}^{x},\ldots,n_{N}^{x})$ in which the value of the $a$-th component $n_{a}^{x}$ denotes the number of particles of species $a\in \{1,\ldots,N\}$. On each vertex we allow a total  maximal occupancy $\nu\in \mathbb{N}$. Then, the configuration space of the process on the graph $G$ is 

@@

with $\tau_{x}\in \{1,\ldots,N\}$, $\forall x\in \{1,\ldots,L\}$. Since the maximal occupancy at each site is $\nu=1$,\todo{but is it at max 1? I do not understand what we call the stirring process....}  the configuration $\bm{n}$ introduced in Section~\ref{subsectionGeneratorStr} and $\bm{\tau}$ are related by 

@@

\subsection{Duality for the integrable process on a line segment}\label{integrableChain-duality}
Here we consider the chain defined in Section~\ref{subsection-description-process-LINE} with the hard-core constrain, i.e. $\nu=1$. 
We recall that the geometry is a one dimensional chain with sites $\{1,\ldots,L\}$ where reservoirs are attached to $1$ and $L$. 
We also recall the notation for the occupation variable which is called $\bm{\tau}=(\tau_1,\ldots,\tau_L)$ with $\tau_x\in\{1,\ldots,N\}$.
The dynamics is described by an Hamiltonian that is written in terms of the first fundamental representation of ${gl}(N)$. The duality result stated in Section~\ref{statementDualitySubsection} is adapted to the present situation as follows. 
The extra-sites are denoted by $0$ and $L+1$, and they are connected with sites $1$ and $L$ of the chain respectively. The matrix $R_{x}$ defined in \eqref{Revmatrix} 
reduces to the identity, therefore the relation between the generators of the algebra and its transposed reduces to
\begin{equation}\label{transpostionPropertyFund}
	e_{ab}^T= e_{ba}\dt
\end{equation}
Therefore, duality matrix reads
\begin{equation}
	D=\mathcal{D}_{0}\otimes\prod_{x=1}^{L}\exp{\left(\sum_{a=1}^{N-1}e_{aN}^{x}\right)}\otimes \mathcal{D}_{L+1}
\end{equation}
where $\mathcal{D}_{0}$ and $\mathcal{D}_{L+1}$ are given in \eqref{dualityMatrix2}. 
The elements of this duality matrix are given by
\begin{equation}\label{ElementsDualityMatrixChain_1}
	D(\bm{\tau},\bm{\xi})=\prod_{a=1}^{N-1}\alpha_{a}^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\prod_{a=1}^{N-1}\mathbbm{1}_{\{\delta_{\tau_{x},a}\geq \xi_{a}^{x}\}}\right)\prod_{a=1}^{N-1}\beta_{a}^{\xi_{a}^{L+1}}\dt
\end{equation}
%\begin{equation}\label{ElementsDualityMatrixChain}
%	D(\bm{n},\bm{\xi})=\prod_{a=1}^{N-1}\alpha_{a}^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\prod_{a=1}^{N-1}\mathbbm{1}_{\{n_{a}^{x}\geq \xi_{a}^{x}\}}\right)\prod_{a=1}^{N-1}\beta_{a}^{\xi_{a}^{L+1}}\dt
%\end{equation}
The dual Hamiltonian is
\begin{equation}
	\widetilde{H}=\widetilde{H}_{\text{left}}+\sum_{x=1}^{L-1}\mathcal{H}_{x,x+1}+\widetilde{H}_{\text{right}}
\end{equation}
where $\mathcal{H}$ is defined in \eqref{H-corsivo} and 
\begin{equation}\label{boundary-H-dual-chain}
	\widetilde{H}_{\text{left}}=\sum_{a=1}^{N-1}\left((\mathbf{a}^{+})_{a}^{0}\,e_{Na}^{1}-e_{aa}^{1}\right)\qquad 	\widetilde{H}_{\text{right}}=\sum_{a=1}^{N-1}\left((\mathbf{a}^{+})_{a}^{L+1}\,e_{Na}^{L}-e_{aa}^{L}\right)\dt
\end{equation}
In the long-time limit the dual process voids the chain, i.e. all particles of types $\{1,\ldots,N-1\}$ are eventually absorbed at the extra-sites $0$ and $L+1$ and replaced by types $N$. We also notice that, up the the bosonic creation operators $(\mathbf{a}^{+})_{a}$, the dual boundary Hamiltonians \eqref{boundary-H-dual-chain} are triangular. This simplification will be crucial in Section~\ref{sectionIntegrabiliy}.\\

@@

{\color{blue} The quantity $\mu(\bm{\tau})$ gives the probability, under the non-equilibrium steady distribution, of having a configuration $\bm{\tau}$, that can be composed by particles and holes as well. However, from Corollary \ref{Corolollary-ABS_Corr-abstract}, we only have an explicit formula for the expectations of the indicator of particle configurations (indeed $a_{1},\ldots,a_{m}$ are different from $N$). Therefore, to write the probability mass function $\mu$ for occupations containing holes we have to use the conservation of probability to get rid of the indicators of these holes. Then, equation \eqref{implicit-mass-function} is derived (for details see below the proof of Corollary \ref{Corollary-NESS-ABS}).\\

@@

	The results in Corollary~\ref{Corolollary-ABS_Corr-abstract} and in Corollary~\ref{Corollary-NESS-ABS}  can be generalized to arbitrary graphs $G=(V,\mathcal{E})$ and for higher maximal occupancies $\nu >1$.

@@

	Thus we chose to present them in such a context. For the chain with higher maximal occupancies $\nu >1$ see Section~\ref{Subsection-ss-nonI}.

@@

In this case, the similarity transformations read

@@

@@

{\color{blue}As a further check we report here the steady state correlations and the vectors $|\Psi^{''}\rangle$, $|\Psi^{'}\rangle$ and $|\Psi\rangle$ for the SSEP, i.e. the situation where the process has only one species of particles and holes. }

@@

In the rest of this section we consider a chain of length $L$, where two reservoirs are attached to sites $1$ and $L$. Each site can host $\nu\geq 1$ particles, whose type is chosen among the $N$ available ones. The graph of the dual process is still a chain, but with two extra-sites $0$ (connected to $1$) and $L+1$ (connected to $L$). The processes considered in this section will be denoted with same name $\left(\bm{n}(t)\right)_{t\geq0}$, their generator will be specified in each subsection. 

@@

where $C(\nu,m)$ is a constant that depends on $m$ and on the maximal occupancy $\nu$. 

@@

	\nu_{TH}(r|\epsilon):&=\mathbb{P}\left(n^{x}=r|n^{x}+n^{x+1}=\epsilon\right)\nonumber

@@

\nu_{TH}(r|\epsilon)&=\frac{\mathbb{P}\left(\left\{n^{x}=r\right\}\,\cap\,\left\{n^{x}+n^{x+1}=\epsilon\right\}\right)}{\mathbb{P}\left(n^{x}+n^{x+1}=\epsilon\right)}=\frac{\mathbb{P}\left(n^{x}=r\right)\,\mathbb{P}\left(n^{x+1}=\epsilon-r\right)}{\mathbb{P}\left(n^{x}+n^{x+1}=\epsilon\right)}\dt

@@

	\nu_{TH}(r|\epsilon)=\frac{\binom{\nu}{r_{1},\ldots,r_{N}}\binom{\nu}{(\epsilon_{1}-r_{1}),\ldots,(\epsilon_{N}-r_{N})}}{\binom{2\nu}{\epsilon_{1}\ldots,\epsilon_{N}}}\mathbbm{1}_{\{r_{1}+\ldots+r_{N}=\nu\}}

@@

	\left\{f(n^{1},\ldots,n^{x-1},r,n^{x}+n^{x+1}-r,n^{x+2},\ldots,n^{L})-f(\bm{n})\right\}\nu_{TH}(\bm{r}|\bm{\epsilon})

@@

In the thermalized bulk generator particles on a bond $x$ and $x+1$ are redistribute according to the measure $\nu_{TH}(r|\epsilon)$, while in the thermalized boundary generators the configuration at site $1$ ($L$) is replaced by $r$, according to a rate depending on $r$ itself and on the reservoirs parameters. 

>>>>

For the single species case,  the solution of the exclusion process with open boundaries \cite{derrida1993exact,1993JSP....72..277S} has been of crucial importance in our understanding of the structural properties of non-equilibrium state, such as long-range correlations \cite{spohn1983long} and non-local density large deviation functions \cite{derrida2007non,derrida1998exact} and current large deviations \cite{mallick2022exact,bodineau2005current}. One may argue that the results available for the exclusion process are rooted in the algebraic symmetries of the corresponding integrable spin $s=1/2$ Heisenberg chain, see e.g. \cite{schutzManyBody,SSEPReviewRagoucy} for two excellent reviews.
%
This integrability property can be combined with duality and leads to closed-form expressions for the correlations and the non-equilibrium steady state, as shown in \cite{frassek2020eigenstates,frassek2020duality}.
% by extending the previous results of \cite{frassek2020eigenstates}.

@@

The model studied in this paper is the {\em  multispecies stirring process with open boundaries}, i.e. the multispecies analogue of the exclusion process on a generic graph as introduced in \cite{zhou2021orthogonal} with additional boundary reservoirs as proposed in \cite{vanicat2017exact}. 
This particle system is defined on a generic connected and undirected graph. At each site 
the maximal number of allowed particles {\color{blue}(also called \textit{maximal occupancy})} is $\nu\in \mathbb{N}$. These particles can be of $N-1$ types (or species). Two sites do interact when there is an edge between them and this interaction consists in swapping the occupations of these two sites (for a precise definition of the process see Section \ref{sec1}). 
Furthermore, reservoirs are attached at each site of the graph, creating and removing particle. The process becomes integrable when the graph is a finite chain with two reservoirs attached to the first and the last sites respectively and the maximal number of allowed particles is fixed to be $\nu=1$,  see \cite{vanicat2017exact}. {\color{blue}The choice of the boundary reservoirs introduced in \cite{vanicat2017exact} is further motivated by the fact that they allow to determine an absorbing dual process, regardless of the geometry on which the process is defined and on the maximal number of particle hosted at each site}.
%{\color{red}Explain the model here.... $N$-species (or $N-1$ plus empty?). What is the graph? say that it becomes integrable for the line }
%We refer to the model as the {\em  multispecies stirring process with open boundaries}.
%

@@

{\color{magenta} Cristian I think it is good to keep some of the following citations, especailly the one to Schutz...
 \cite{schutz2017kardar, crampe2016integrable, finn2018matrix, floreani2022switching, casini2023density, redig2022ergodic, redig2023equilibrium, chatterjee2023multi,  franceschini2022orthogonal,floreani2023non,carinci2023open}
See also \cite{crampe2016integrable, finn2018matrix} for the asymmetric setup.}

@@

At each vertex site $x\in V$, we describe the occupation with an $N-$dimensional vector $n^{x}=(n_{1}^{x},\ldots,n_{N}^{x})$ in which the value of the $a$-th component $n_{a}^{x}$ {\color{blue} denotes the number of particles of species $a\in \{1,\ldots,N-1\}$, while the component $n_{N}^{x}$ denotes the number of holes at site $x$.} On each vertex we {\color{blue} allow a total maximal number of particles $\nu\in \mathbb{N}$}. Then, the configuration space of the process on the graph $G$ is 

@@

with $\tau_{x}\in \{1,\ldots,N\}$, $\forall x\in \{1,\ldots,L\}$. Since the {\color{blue} maximal occupancy at each site is $\nu=1$},\todo{but is it at max 1? I do not understand what we call the stirring process....}  the configuration $\bm{n}$ introduced in Section~\ref{subsectionGeneratorStr} and $\bm{\tau}$ are related by 

@@

\subsection{Duality for the integrable process on a line segment}\label{integrableChain-duality}
Here we consider the chain defined in Section~\ref{subsection-description-process-LINE} with the hard-core constrain {\color{blue} (at most one particle per site)}, i.e. $\nu=1$. 
We recall that the geometry is a one dimensional chain with sites $\{1,\ldots,L\}$ where reservoirs are attached to $1$ and $L$. 
We also recall the notation for the occupation variable which is called $\bm{\tau}=(\tau_1,\ldots,\tau_L)$ with $\tau_x\in\{1,\ldots,N\}$.
The dynamics is described by an Hamiltonian that is written in terms of the first fundamental representation of ${gl}(N)$. The duality result stated in Section~\ref{statementDualitySubsection} is adapted to the present situation as follows. 
The extra-sites are denoted by $0$ and $L+1$, and they are connected with sites $1$ and $L$ of the chain respectively. The matrix $R_{x}$ defined in \eqref{Revmatrix} 
reduces to the identity, therefore the relation between the generators of the algebra and its transposed reduces to
\begin{equation}\label{transpostionPropertyFund}
	e_{ab}^T= e_{ba}\dt
\end{equation}
Therefore, duality matrix reads
\begin{equation}
	D=\mathcal{D}_{0}\otimes\prod_{x=1}^{L}\exp{\left(\sum_{a=1}^{N-1}e_{aN}^{x}\right)}\otimes \mathcal{D}_{L+1}
\end{equation}
where $\mathcal{D}_{0}$ and $\mathcal{D}_{L+1}$ are given in \eqref{dualityMatrix2}. 
The elements of this duality matrix are given by
\begin{equation}\label{ElementsDualityMatrixChain_1}
	D(\bm{\tau},\bm{\xi})=\prod_{a=1}^{N-1}\alpha_{a}^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\prod_{a=1}^{N-1}\mathbbm{1}_{\{\delta_{\tau_{x},a}\geq \xi_{a}^{x}\}}\right)\prod_{a=1}^{N-1}\beta_{a}^{\xi_{a}^{L+1}}\dt
\end{equation}
%\begin{equation}\label{ElementsDualityMatrixChain}
%	D(\bm{n},\bm{\xi})=\prod_{a=1}^{N-1}\alpha_{a}^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\prod_{a=1}^{N-1}\mathbbm{1}_{\{n_{a}^{x}\geq \xi_{a}^{x}\}}\right)\prod_{a=1}^{N-1}\beta_{a}^{\xi_{a}^{L+1}}\dt
%\end{equation}
The dual Hamiltonian is
\begin{equation}
	\widetilde{H}=\widetilde{H}_{\text{left}}+\sum_{x=1}^{L-1}\mathcal{H}_{x,x+1}+\widetilde{H}_{\text{right}}
\end{equation}
where $\mathcal{H}$ is defined in \eqref{H-corsivo} and 
\begin{equation}\label{boundary-H-dual-chain}
	\widetilde{H}_{\text{left}}=\sum_{a=1}^{N-1}\left((\mathbf{a}^{+})_{a}^{0}\,e_{Na}^{1}-e_{aa}^{1}\right)\qquad 	\widetilde{H}_{\text{right}}=\sum_{a=1}^{N-1}\left((\mathbf{a}^{+})_{a}^{L+1}\,e_{Na}^{L}-e_{aa}^{L}\right)\dt
\end{equation}
In the long-time limit the dual process voids the chain, i.e. all particles of types $\{1,\ldots,N-1\}$ are eventually absorbed at the extra-sites $0$ and $L+1$ and replaced by types $N$. We also notice that, up the the bosonic creation operators $(\mathbf{a}^{+})_{a}$, the dual boundary Hamiltonians \eqref{boundary-H-dual-chain} are triangular. This simplification will be crucial in Section~\ref{sectionIntegrabiliy}.\\

@@

{\color{blue} The quantity $\mu(\bm{\tau})$ is the probability, under the non-equilibrium steady distribution, of having a configuration $\bm{\tau}$, that can be composed by particles and holes as well. However, from Corollary \ref{Corolollary-ABS_Corr-abstract}, we only have an explicit formula for the expectations of the indicator of particle configurations (indeed $a_{1},\ldots,a_{m}$ are different from $N$). Therefore, to write the probability mass function $\mu$ for occupations containing holes we have to use the conservation of probability to get rid of the indicators of these holes. Then, equation \eqref{implicit-mass-function} is derived (for details see below the proof of Corollary \ref{Corollary-NESS-ABS}).\\

@@

	The results in Corollary~\ref{Corolollary-ABS_Corr-abstract} and in Corollary~\ref{Corollary-NESS-ABS}  can be generalized to arbitrary graphs $G=(V,\mathcal{E})$ and for {\color{blue} higher maximal number of particles allowed at each site}, that is for $\nu >1$.

@@

	Thus we chose to present them in such a context. For the chain with {\color{blue} more than one particle allowed at each site} $\nu >1$ see Section~\ref{Subsection-ss-nonI}.

@@

where we assume that $(\alpha_{1}+\alpha_{2})=(\beta_{1}+\beta_{2})=1$. In this case, the similarity transformations read

@@

\end{pmatrix},\qquad \mathcal{S}=\mathcal{S}_{2}\mathcal{S}_{1}=\begin{pmatrix}
1-\beta_{1}&-\beta_{1}\\

@@

{\color{blue}As a further check we report the steady state correlations and the vectors $|\Psi^{''}\rangle$, $|\Psi^{'}\rangle$ and $|\Psi\rangle$ for the SSEP, i.e. the situation where the process has only one species of particles and holes, with maximal occupancy $\nu=1$. }

@@

In the rest of this section we consider a chain of length $L$, where two reservoirs are attached to sites $1$ and $L$. Each site can {\color{blue} host at most $\nu\geq 1$ particles per site}, whose type is chosen among the $N$ available ones. The graph of the dual process is still a chain, but with two extra-sites $0$ (connected to $1$) and $L+1$ (connected to $L$). The processes considered in this section will be denoted with same name $\left(\bm{n}(t)\right)_{t\geq0}$, their generator will be specified in each subsection. 

@@

where $C(\nu,m)$ is a constant that depends on $m$ and on the maximal number of particle per site $\nu$. 

@@

	\mu_{TH}(r|\epsilon):&=\mathbb{P}\left(n^{x}=r|n^{x}+n^{x+1}=\epsilon\right)\nonumber

@@

\mu_{TH}(r|\epsilon)&=\frac{\mathbb{P}\left(\left\{n^{x}=r\right\}\,\cap\,\left\{n^{x}+n^{x+1}=\epsilon\right\}\right)}{\mathbb{P}\left(n^{x}+n^{x+1}=\epsilon\right)}=\frac{\mathbb{P}\left(n^{x}=r\right)\,\mathbb{P}\left(n^{x+1}=\epsilon-r\right)}{\mathbb{P}\left(n^{x}+n^{x+1}=\epsilon\right)}\dt

@@

	\mu_{TH}(r|\epsilon)=\frac{\binom{\nu}{r_{1},\ldots,r_{N}}\binom{\nu}{(\epsilon_{1}-r_{1}),\ldots,(\epsilon_{N}-r_{N})}}{\binom{2\nu}{\epsilon_{1}\ldots,\epsilon_{N}}}\mathbbm{1}_{\{r_{1}+\ldots+r_{N}=\nu\}}

@@

	\left\{f(n^{1},\ldots,n^{x-1},r,n^{x}+n^{x+1}-r,n^{x+2},\ldots,n^{L})-f(\bm{n})\right\}\mu_{TH}(\bm{r}|\bm{\epsilon})

@@

In the thermalized bulk generator particles on a bond $x$ and $x+1$ are redistribute according to the measure $\mu_{TH}(r|\epsilon)$, while in the thermalized boundary generators the configuration at site $1$ ($L$) is replaced by $r$, according to a rate depending on $r$ itself and on the reservoirs parameters. 
